A Type System for Safe and Scalable Parallelism using Logical Regions

### 2012-07-07 doc/popl2013/langdef.tex
new sections

\subsection{Soundness of Effects}

@@

\subsection{Coherence}
\label{sec:coherence}

Theorem~\ref{thm:effects} guarantees expression evaluation produces a value of the
correct type and does so within the regions of memory for which privileges are held.

 but for the
Legion programming model to be useful, we need to also guarantee that the evaluation of an expression
produces the (or at least, a) correct value.  The key to doing this is to place additional 
constraints on what it means to be a valid interleaving of two or more memory traces.  In
addition to a valid interleaving being an interleaving of the individual traces, we will 
require that it be:
\begin{itemize}
\item {\em coherent} - each memory read operation should see the effect of all writes and reductions 
that occur earlier in the interleaved trace to the same location, and
\item {\em sequentially equivalent} - the value read by each memory read operation and the result of applying the interleaved trace to the original store should be the same as if each constituent trace
was applied to the store in a valid {\em program order}.
\end{itemize}
These requirements are relaxed in some cases due to either a non-empty clobber set (representing
operations that may occur concurrently as part of evaluating other functions) or due to coherence
modes other than {\em excl} on operations with the traces (representing operations that may 
occur concurrently as part of evaluating other parts of the same function).

@@

\subsection{Hierarchical Scheduling}

>>>>

\section{Soundness of Effects}

@@

\section{Coherence}
\label{sec:coherence}

Theorem~\ref{thm:effects} guarantees that a task produces a value of the
correct type and does so within the regions of memory for which privileges are held.
It does not, however, address conherence, what concurrent operations on the regions are permitted.
An innovation in Legion is that programmers can control coherence on a per-region basis.  Intuitively,
the different coherence modes permit different interleavings of memory traces.  We begin by formalizing
exclusive access, the strictest coherence mode.  Exclusive access requires valid interleavings be:
\begin{itemize}
\item {\em coherent} - a read of location $l$ should see the effect of all writes and reductions to $l$
occuring earlier in the interleaved trace, and
\item {\em sequentially equivalent} - the value read by each read 
operation and the result of applying the interleaved trace to the original store should be the same as if each constituent trace
was applied to the store in a valid sequential execution of the program.
\end{itemize}
These requirements are relaxed in some cases due to either a non-empty clobber set (representing
operations that may occur concurrently as part of evaluating other functions) or for coherence
modes other than {\em excl}.

@@

\section{Hierarchical Scheduling}
\label{sec:scheduling}
